---
title: "Clinical Tissue Indices Explained"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Clinical Tissue Indices Explained}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CTTIR/hyperspectR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CTTIR/hyperspectR/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/CTTIR/hyperspectR/actions/workflows/pkgdown.yaml/badge.svg)](https://cttir.github.io/hyperspectR/)
[![CRAN status](https://www.r-pkg.org/badges/version/hyperspectR)](https://CRAN.R-project.org/package=hyperspectR)
[![Codecov test coverage](https://codecov.io/gh/CTTIR/hyperspectR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/CTTIR/hyperspectR?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/hyperspectR)](https://cran.r-project.org/package=hyperspectR)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/hyperspectR)](https://cran.r-project.org/package=hyperspectR)
[![License: MIT](https://img.shields.io/badge/license-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(hyperspectR)
```

## Hemoglobin Absorption

Tissue oxygenation assessment via HSI relies on the distinct absorption
spectra of oxyhemoglobin (HbO2) and deoxyhemoglobin (Hb).

```{r}
hb_data <- hs_chromophore_data(c("HbO2", "Hb"), wavelength_range = c(430, 910))
library(ggplot2)
ggplot(hb_data, aes(x = wavelength)) +
  geom_line(aes(y = HbO2, color = "HbO2"), linewidth = 0.8) +
  geom_line(aes(y = Hb, color = "Hb"), linewidth = 0.8) +
  scale_color_manual(values = c(HbO2 = "#E41A1C", Hb = "#377EB8")) +
  labs(x = "Wavelength (nm)", y = "Extinction Coefficient",
       title = "Hemoglobin Absorption Spectra", color = "") +
  theme_hsi()
```

## Tissue Oxygen Saturation (StO2)

StO2 estimates superficial tissue oxygenation (~1 mm depth) by comparing
reflectance in the visible Hb absorption region (500-650 nm) to the NIR
region (700-815 nm).

```{r}
cube <- hs_example_cube()
sto2 <- hs_sto2(cube)
hs_plot_index(sto2, title = "StO2 (%)", palette = "sto2")
```

## Near-Infrared Perfusion Index (NPI)

NPI probes deeper tissue layers (4-6 mm) using NIR wavelengths (655-910 nm).

```{r}
npi <- hs_npi(cube)
hs_plot_index(npi, title = "NPI (%)", palette = "perfusion")
```

## Tissue Hemoglobin Index (THI)

THI estimates relative hemoglobin concentration from the Q-band absorption
depth (530-590 nm) relative to a reference region (785-825 nm).

```{r}
thi <- hs_thi(cube)
hs_plot_index(thi, title = "THI (%)", palette = "hemoglobin")
```

## Custom Normalized Difference Index

Create any two-band ratio index:

```{r}
ndi <- hs_ndi(cube, band1 = 540, band2 = 660)
hs_plot_index(ndi, title = "NDI (540/660)", range = c(-1, 1))
```

## Clinical Panel

The TIVITA-style 5-panel display combines RGB with all indices:

```{r fig.width=10, fig.height=4}
hs_plot_clinical(cube)
```
