---
title: "Getting Started with hyperspectR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with hyperspectR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CTTIR/hyperspectR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CTTIR/hyperspectR/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/CTTIR/hyperspectR/actions/workflows/pkgdown.yaml/badge.svg)](https://cttir.github.io/hyperspectR/)
[![CRAN status](https://www.r-pkg.org/badges/version/hyperspectR)](https://CRAN.R-project.org/package=hyperspectR)
[![Codecov test coverage](https://codecov.io/gh/CTTIR/hyperspectR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/CTTIR/hyperspectR?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/hyperspectR)](https://cran.r-project.org/package=hyperspectR)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/hyperspectR)](https://cran.r-project.org/package=hyperspectR)
[![License: MIT](https://img.shields.io/badge/license-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Installation

```{r eval=FALSE}
# Install from GitHub
# install.packages("remotes")
remotes::install_github("cttir/hyperspectR")
```

## Loading the Package

```{r setup}
library(hyperspectR)
```

## Creating and Inspecting a Cube

hyperspectR provides synthetic data generators for testing and demonstration.
`hs_example_cube()` creates a 30x30 pixel tissue scene with 61 spectral bands
spanning 430-910 nm:

```{r}
cube <- hs_example_cube()
print(cube)
```

Check the dimensions:

```{r}
dim(cube)
```

The summary provides per-band statistics:

```{r}
s <- summary(cube)
s$wavelength_range
s$data_range
```

## Subsetting

You can subset cubes spatially and spectrally while preserving the class:

```{r}
sub <- cube[1:15, 1:15, 1:30]
dim(sub)
```

## Visualization

### RGB Composite

```{r}
ggplot2::autoplot(cube, type = "rgb")
```

### Single Band Image

```{r}
hs_plot_image(cube, wavelength = 550)
```

### Mean Spectrum

```{r}
hs_plot_spectra(cube, pixels = "mean", show_sd = TRUE)
```

## Computing Tissue Indices

The package implements clinical tissue indices used in surgical HSI:

```{r}
sto2 <- hs_sto2(cube)
hs_plot_index(sto2, title = "StO2 (%)", palette = "sto2")
```

### Clinical Panel Display

```{r fig.width=10, fig.height=4}
hs_plot_clinical(cube, indices = c("sto2", "npi", "thi"))
```

## Interactive Exploration

Launch the Shiny app for interactive analysis:

```{r eval=FALSE}
hs_run_app(cube)
```

## Next Steps

- Read real data with `hs_read_envi()` or `hs_read_cube()`
- Preprocess with `hs_smooth()`, `hs_snv()`, `hs_msc()`
- Perform PCA with `hs_pca()` or MNF with `hs_mnf()`
- Classify with `hs_sam()` or fit chromophores with `hs_beer_lambert()`

# Use of LLM tools

Portions of this package were prepared with assistance from large language model tooling for
narrowly defined, non-authorial tasks: copyediting, prose smoothing, Markdown/LaTeX formatting,
scaffolding of boilerplate files (CI configs, build scripts), code refactoring. The tools used were [Chat AI](https://kisski.gwdg.de/leistungen/2-02-llm-service/),
the LLM service of KISSKI (GWDG), and a self-hosted **Mistral Small (24B, Apache-2.0)** run locally via
[Ollama](https://ollama.com/) and the `ollamar` R package — local inference only, with no data sent to
third parties for the self-hosted model.

All scientific claims, methodological choices, analyses, interpretations, and conclusions are the
author's own. No LLM-generated text was incorporated without review and revision, and every reference
was verified against its DOI, arXiv ID, or ISBN.
