---
title: "Research Oxygenation Model Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Research Oxygenation Model Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- badges: start -->
[![R-CMD-check](https://github.com/CTTIR/hyperspectR/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/CTTIR/hyperspectR/actions/workflows/R-CMD-check.yaml)
[![pkgdown](https://github.com/CTTIR/hyperspectR/actions/workflows/pkgdown.yaml/badge.svg)](https://cttir.github.io/hyperspectR/)
[![CRAN status](https://www.r-pkg.org/badges/version/hyperspectR)](https://CRAN.R-project.org/package=hyperspectR)
[![Codecov test coverage](https://codecov.io/gh/CTTIR/hyperspectR/branch/main/graph/badge.svg)](https://app.codecov.io/gh/CTTIR/hyperspectR?branch=main)
[![CRAN downloads](https://cranlogs.r-pkg.org/badges/hyperspectR)](https://cran.r-project.org/package=hyperspectR)
[![CRAN downloads total](https://cranlogs.r-pkg.org/badges/grand-total/hyperspectR)](https://cran.r-project.org/package=hyperspectR)
[![License: MIT](https://img.shields.io/badge/license-MIT-yellow.svg)](https://opensource.org/licenses/MIT)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

```{r setup}
library(hyperspectR)
```

## Overview

This vignette demonstrates a research processing workflow on a synthetic scene.
The simulator is not a tissue measurement or an independent validation dataset.
No surgical decision or diagnostic threshold follows from these examples.

## Step 1: Load Data

```{r}
cube <- hs_simulate_cube(rows = 50, cols = 50, n_regions = 4,
                          sto2_range = c(0.2, 0.95), seed = 42)
print(cube)
```

## Step 2: Preprocessing

```{r}
smoothed <- hs_smooth(cube, window = 7, poly = 3)
hs_plot_spectra(smoothed)
```

## Step 3: Dimensionless Band Ratio

```{r}
sto2_ratio <- hs_band_ratio(smoothed, c(700, 800), c(500, 650))
hs_plot_index(sto2_ratio, title = "Relative NIR/visible ratio", palette = "sto2")
```

## Step 4: Beer-Lambert Chromophore Fitting

```{r}
fit <- hs_beer_lambert(smoothed)
hs_plot_index(fit$sto2, title = "Fitted Hb fraction (%)", palette = "sto2")
```

## Step 5: ROI-Based Statistical Analysis

```{r}
# Define ROIs for different surgical field regions
roi_healthy <- hs_roi_rect(smoothed, x_range = c(1, 20), y_range = c(1, 20))
roi_ischemic <- hs_roi_rect(smoothed, x_range = c(30, 50), y_range = c(1, 20))

stats_healthy <- hs_roi_stats(smoothed, roi_healthy)
stats_ischemic <- hs_roi_stats(smoothed, roi_ischemic)

library(ggplot2)
ggplot() +
  geom_ribbon(data = stats_healthy,
              aes(x = wavelength, ymin = mean - sd, ymax = mean + sd),
              fill = "#2E86AB", alpha = 0.3) +
  geom_line(data = stats_healthy,
            aes(x = wavelength, y = mean, color = "Region 1"), linewidth = 0.8) +
  geom_ribbon(data = stats_ischemic,
              aes(x = wavelength, ymin = mean - sd, ymax = mean + sd),
              fill = "#E41A1C", alpha = 0.3) +
  geom_line(data = stats_ischemic,
            aes(x = wavelength, y = mean, color = "Region 2"), linewidth = 0.8) +
  scale_color_manual(values = c("Region 1" = "#2E86AB", "Region 2" = "#E41A1C")) +
  labs(x = "Wavelength (nm)", y = "Reflectance",
       title = "ROI Spectral Comparison", color = "") +
  theme_hsi()
```

## Step 6: Research Panel

```{r fig.width=10, fig.height=4}
hs_plot_clinical(smoothed)
```

## Step 7: PCA for Quality Assessment

```{r fig.width=10, fig.height=4}
pca <- hs_pca(smoothed, n_components = 3)
cat("Variance explained:", round(pca$variance_explained * 100, 1), "%\n")
```
